---
title: "Lab2"
author: "Bhavya Patel"
date: "2024-09-15"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2movies) 
data(movies)
library(dplyr)
library(tidyr)
library(ggplot2) 
```


##Q1
```{r Q1}
a=range(movies$year)
paste("The oldest movie was in ",a[1],sep=" ")
paste("This latest movie was in",a[2],sep=" ")
```

The Range of years of production of the movies is from 1893 to 2005 according to
this dataset.

##Q2
```{r Q2}
bn=as.integer(table(is.na(movies$budget))[1])
bm=as.integer(table(is.na(movies$budget))[2])
p1=bn/(bn+bm)
p1=p1*100
p2=bm/(bn+bm)
p2=p2*100
paste("This is the proportion of movies that have their budget included in this 
      database", round(p2,2),"%",sep=" ")
paste("This is the proportion of movies that donot have their budget included in 
      this database", round(p1,2),"%",sep=" ")
print("Here's the lsit top 5 expensive movies")
mb <- movies %>% 
  select(title, budget) %>% 
  arrange(desc(budget)) %>% 
  head(5)

mb[,1]
```

The proportion of the movies that included their budget is given above and the 
top 5 expensive movies list is also provided above.

##Q3
```{r Q3}
ml<- movies %>%
  select(title,length) %>%
  arrange(desc(length)) %>%
  head(5)
ml

```

The list of top 5 longest movies is provided above.

##Q4
```{r Q4}
ss <- movies %>%
  filter(Short==1) %>%
  select(title,length) %>%
  arrange(length)  %>%
  head(1)
paste("The shorted Short film is",ss[1],sep=" ")

ls <- movies %>%
  filter(Short==1) %>%
  select(title,length) %>%
  arrange(desc(length))  %>%
  head(1)
paste("The longest Short film is", ls[1], sep=" ")

```

The shortest and longest Short film list is provided above.

##Q5
```{r Q5}
gc <- movies %>%
  pivot_longer(cols = Action:Short, names_to = "gc", values_to = "present") %>%
  filter(present == 1) %>%   
  group_by(gc) %>%
  summarise(count = n())

ggplot(gc, aes(x = gc, y = count, fill = gc)) + geom_bar(stat = "identity") +
labs(title = "Number of Movies by Genre", x = "Genre", y = "Number of Movies") +
scale_fill_manual(values = c("Action" = "red", "Animation" = "blue", 
    "Comedy" = "green", "Drama" = "purple", "Documentary" = "orange", 
    "Romance" = "pink", "Short" = "yellow")) + 
  guides(fill=guide_legend(title="Genres"))
```

This bar chart represents number of movies in each Genre.

##Q6
```{r Q6}
gr <- movies %>%
  pivot_longer(cols = Action:Short, names_to = "gc", values_to = "present") %>%
  filter(present == 1) %>%   
  group_by(gc) %>%
  summarise(avg = mean(rating, na.rm = TRUE))

ggplot(gr, aes(x = gc, y = avg, fill = gc)) + geom_bar(stat = "identity") +
  labs(title = "Average Rating of Movies by Genre", x = "Genre", y = "Ratings")+
  scale_fill_manual(values = c("Action" = "red", "Animation" = "blue", 
  "Comedy" = "green", "Drama" = "purple", "Documentary" = "orange", 
  "Romance" = "pink", "Short" = "yellow")) + 
  guides(fill=guide_legend(title="Genres"))
```

This bar chart represents average rating of movies in each Genre.

##Q7
```{r Q7}
gr2 <- movies %>%
  pivot_longer(cols = Action:Short, names_to = "gc", values_to = "present") %>%
  filter(present == 1, year >= 2000 & year <= 2005) %>%   
  group_by(gc) %>%
  summarise(avg = mean(rating, na.rm = TRUE))

ggplot(gr2, aes(x = gc, y = avg, fill = gc)) + geom_bar(stat = "identity") +
  labs(title = "Average Rating of Movies by Genre Between 2000-2005", x = "Genre", y = "Ratings") +
  scale_fill_manual(values = c("Action" = "red", "Animation" = "blue", 
  "Comedy" = "green", "Drama" = "purple", "Documentary" = "orange", 
  "Romance" = "pink", "Short" = "yellow"))+ 
  guides(fill=guide_legend(title="Genres"))

```

This bar chart represents average rating of movies in each Genre between 
2000-2005.

##Q8
```{r Q8}
gm <- movies %>%
  pivot_longer(cols = Action:Romance, names_to = "gc", values_to = "present") %>%  
  filter(present == 1 & year >= 1990) %>%
  group_by(year, gc) %>%
  summarise(count = n(), .groups = "drop") 

ggplot(gm, aes(x = year, y = count, color = gc)) + geom_line(size = 1.5) +
  labs(title = "Number of Movies by Genre Produced Per Year (1990 onwards)", 
       x = "Genre", y = "Number of Movies", color="Genre") +
  scale_color_manual(values = c("Action" = "red", "Animation" = "blue", 
    "Comedy" = "green", "Drama" = "purple", "Documentary" = "orange", 
    "Romance" = "pink"))
```

This line chart represents the number of movies produced yearly in each Genre 
except Short movies from 1990 onwards.  

##Q9
```{r Q9}
paste("Question 1. What are the top 10 R-rated movies with best ratings?")
tr <- movies %>%
  filter(mpaa=="R") %>%
  arrange(desc(rating)) %>%
  select(title, rating) %>%
  head(10)

#Answer 1:
paste("Top 10 best R-rated movies:")
tr[,1]

paste("Question 2. What movie has the best ratings in Action and Drama?") 
paste("How much was their budget, and what was the length of the movie?")
bad <- movies %>%
  filter(Action==1 & Drama==1) %>%
  arrange(desc(rating)) %>%
  select(title, rating, budget, length) %>%
  head(1)

#Answer2:
paste("The Best Rated movie in Action and Drama is", bad$title, 
      "with ratings of", bad$rating, sep=" ")
paste("Their budge was", bad$budget, "and the length of the movies was", 
      bad$length, "minutes.", sep=" ")


paste("Question 3. What is the average budget for Action films over the past") 
paste("20 years (use a line chart)?")
t1=max(movies$year)-20
aba <- movies %>%
  filter(year>t1, Action==1) %>%
  group_by(year) %>%
  summarise(avg = mean(budget, na.rm=T))

aba$avg <- aba$avg/1e6

ggplot(aba, aes(x = year, y = avg)) +
  geom_line(color = "blue", size = 1.5) +
  geom_point(color = "red", size = 3) +
  labs(title = "Average Budget of Action Films (Last 20 Years)",
       x = "Year of Release",
       y = "Average Budget in Millions")

```

Here I have added 3 of my own questions and answered them as well.